---
title: "Supplementary Figure 4"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
    includes:
      in_header: header.tex
  html_document:
    fig_caption: yes
    self_contained: no
    toc: yes
---

```{r setup, include=FALSE}
fn = local({
  function(x, letter) {
    paste('__Supplementary Figure 4', letter, '__: ', x, sep = '')
  }
})
```


```{r loadLib, message=FALSE, warning=FALSE, error=FALSE, echo=FALSE}
##load the library
library(motifStack)
colorSet <- c("Dm"="#00FC00", "b1h"="#00FC00", "sw"="#008080", "bml"="darkgreen", 
              "Mm"="brown", "MmDREAM"="blue", "Ms"="#F69156", 
              "Hs"="#D900D9")

## function to read example data
getMatAlignOut <- function(pcmpath, outpath="output", groupDistance=2.5, trim=0.2){
    pcms <- readPCM(pcmpath)
    pfms<-lapply(pcms,pcm2pfm)
    matalign_path <- "./app/matalign-v4a"
    neighbor_path <- "./app/neighbor.app/Contents/MacOS/neighbor"
    system(paste("perl MatAlign2tree.pl --in . --pcmpath", pcmpath, "--out", outpath,
             "--matalign", matalign_path, "--neighbor", neighbor_path, "--tree","UPGMA"))
    newickstrUPGMA <- readLines(con=file.path(outpath, "NJ.matalign.distMX.nwk"))
    phylog <- newick2phylog(newickstrUPGMA, FALSE)
    leaves <- names(phylog$leaves)
    motifs <- pfms[leaves]
    if(!is.na(groupDistance)){
        motifSig <- motifSignature(motifs, phylog, groupDistance=groupDistance,
                                   min.freq=1, trim=trim)
        sig <- signatures(motifSig)
        gpCol <- sigColor(motifSig)
    }else{
        motifSig <- NA
        sig <- NA
        gpCol <- NA
    }
    
    return(list(phylog=phylog, sig=sig, gpCol=gpCol, 
                motifs=DNAmotifAlignment(motifs), leaves=leaves, 
                unaligned.pfms=motifs))
}

pfmList2matrixList <- function(pfms){
        m <- lapply(pfms, function(.ele) as(.ele, "matrix"))
        names(m) <- unlist(lapply(pfms, function(.ele) .ele@name))
        m
}

getMotIVOut <- function(pcmpath, cc, align, groupDistance=.005, trim=0.2){
    pcms <- readPCM(pcmpath)
    pfms<-lapply(pcms,pcm2pfm)
    jaspar.scores <- 
        MotIV::readDBScores(
            file.path(".", "app", "scores", 
                      paste("JaspRand_", cc, "_", align, ".scores", sep="")))
    d <- MotIV::motifDistances(pfmList2matrixList(pfms), cc=cc, align=align)
    hc <- MotIV::motifHclust(d)
    phylog <- hclust2phylog(hc)
    pfms <- pfms[hc$order]
    aligned.pfms <- DNAmotifAlignment(pfms)
    leaveNames <-names(phylog$leaves)
    leaves <- names(phylog$leaves)
    motifs <- pfms[leaves]
    if(!is.na(groupDistance)){
        motifSig <- motifSignature(motifs, phylog, groupDistance=groupDistance,
                                   min.freq=1, trim=trim)
        sig <- signatures(motifSig)
        gpCol <- sigColor(motifSig)
    }else{
        motifSig <- NA
        sig <- NA
        gpCol <- NA
    }
    
    return(list(phylog=phylog, sig=sig, gpCol=gpCol, motifs=DNAmotifAlignment(motifs),
                leaves=leaves, unaligned.pfms=motifs))
}

```

# Supplementary Figure 4A. Align motifs from fly by MatAlign
```{r Fly_B1H, fig.width=8, fig.height=11, dpi=72, fig.cap=fn("Alignment of fly motifs using MatAlign", "A"), echo=FALSE}
gpDis <- 2
Fly_B1H <- getMatAlignOut("pcmsDatasetFly", groupDistance=gpDis)
attach(Fly_B1H)

leaveNames <- gsub("^Dm_", "", leaves)

motifPiles(phylog=phylog, motifs, sig, 
           col.pfms=gpCol, col.pfms.width=.01, 
           col.pfms2=gpCol, col.pfms2.width=.01, 
           labels.leaves=leaveNames, 
           plotIndex=c(FALSE, TRUE), IndexCex=1, 
           groupDistance=gpDis, clabel.leaves=1)

detach(Fly_B1H)
```

# Supplementary Figure 4B. Alignment motifs from fly using MotIV

```{r Fly_B1H.MotIV, fig.width=8, fig.height=11, dpi=72, fig.cap=fn("Alignment of fly motifs using MotIV", "B"), echo=FALSE}
gpDis <- 0.005 
Fly_B1H.MotIV <- getMotIVOut("pcmsDatasetFly", 
                                    cc="ALLR", align="SWU", 
                                    groupDistance=gpDis)
attach(Fly_B1H.MotIV)

leaveNames <- gsub("^Dm_", "", leaves)
motifPiles(phylog=phylog, motifs, sig, 
           col.pfms=gpCol, col.pfms.width=.01, 
           col.pfms2=gpCol, col.pfms2.width=.01, 
           labels.leaves=leaveNames, 
           plotIndex=c(FALSE, TRUE), IndexCex=1, 
           groupDistance=gpDis, clabel.leaves=1)

detach(Fly_B1H.MotIV)
```

# Supplementary Figure 4C. Align motifs from fly and mouse using MatAlign
```{r FlyMousePBMSAW, fig.width=12, fig.height=12, dpi=72, fig.cap=fn("Alignment of motifs of fly and mouse using MatAlign", "C"), echo=FALSE}
gpDis <- 2
FlyMousePBMSAW <- getMatAlignOut("pcmsDatasetFlyMousePBM.SAW", 
                                 groupDistance=gpDis, trim=0.2)
attach(FlyMousePBMSAW)

##set methods color
species <- gsub("^(Dm|Mm)_.*$", "\\1", leaves)
species <- colorSet[species]##c("Dm/FFS/B1H", "Mm/Uniprobe/PBM/Seed-And-Wobble")

leaveNames <- gsub("^(Dm|Mm)_", "", leaves)

plotMotifStackWithRadialPhylog(phylog=phylog, pfms=sig, 
                               col.bg=species, col.bg.alpha=.3, 
                               col.inner.label.circle=gpCol, 
                               inner.label.circle.width=0.02, 
                               labels.leaves=leaveNames, 
                               cleaves=.1, circle=1.35, circle.motif=1.6, 
                               clabel.leaves=.5, motifScale="logarithmic", 
                               angle=358, plotIndex=TRUE, IndexCex=.6, 
                               groupDistance=gpDis)

legend(-2.2, 2.4, legend=c("Dm/FFS/B1H", "Mm/Uniprobe/PBM/\nSeed-And-Wobble"), 
       fill= highlightCol(colorSet[c("Dm", "Mm")], alpha=.3), 
       border="white", lty=NULL, bty = "n")

detach(FlyMousePBMSAW)
```

# Supplementary Figure 4D. Align motifs from fly and human using MatAlign
```{r FlyHuman, fig.width=12, fig.height=12, dpi=72, fig.cap=fn("Alignment of motifs from fly and human using MatAlign", "D"), echo=FALSE}
gpDis <- 2
FlyHuman <- getMatAlignOut("pcmsDatasetFlyHuman", groupDistance=gpDis)
attach(FlyHuman)

##set methods color
species <- gsub("^(Dm|Hs)_.*$", "\\1", leaves)
species <- colorSet[species] ##c("Dm/FFS/B1H", "Hs/CIS-BP/SELEX")

leaveNames <- gsub("^(Dm|Hs)_", "", leaves)

plotMotifStackWithRadialPhylog(phylog=phylog, pfms=sig, col.bg=species, 
                               col.bg.alpha=.3, col.inner.label.circle=gpCol, 
                               inner.label.circle.width=0.04, 
                               labels.leaves=leaveNames, 
                               cleaves=.1, circle=1.35, circle.motif=1.8, 
                               clabel.leaves=.5, motifScale="logarithmic", 
                               angle=358, plotIndex=TRUE, IndexCex=.6, 
                               groupDistance=gpDis)

legend(-2, 2.2, legend=c("Dm/FFS/B1H", "Hs/CIS-BP/SELEX"), 
       fill=highlightCol(colorSet[c("Dm", "Hs")], alpha=.3), 
       border="white", lty=NULL, bty = "n")

detach(FlyHuman)
```


# Supplementary Figure 4E. Align motifs from mouse generated by different methods (SELEX, PBM-SW and PBM-dream5) using MatAlign
```{r MmPBM_SELEX, fig.width=12, fig.height=12, dpi=72, fig.cap=fn("Alignment of ouse data generated by different methods using MatAlign", "E.a")}
gpDis <- 2
MmPBM_SELEX <- getMatAlignOut("pcmsDatasetMmPBM_SELEX", groupDistance=gpDis)
attach(MmPBM_SELEX)

##set methods color
methods <- factor(grepl("_SELEX", leaves) + grepl("_M\\d{4}", leaves))
levels(methods) <- c("Mm", "MmDREAM", "Ms")
##c("PBM/Uniprobe/Seed-And-Wobble", "PBM/CIS-BP/DREAM5", "SELEX/CIS-BP")
levels(methods) <- colorSet[levels(methods)]
methods <- as.character(methods)

leaveNames <- gsub("_(SELEX|PBM)", "", gsub("_(Jolma|Berger)", "", leaves))

## calculate average of top 8 position information content for each motif
icgp <- sapply(sapply(unaligned.pfms, getIC), function(.ele) 
  mean(sort(.ele, decreasing=TRUE)[1:min(8, length(.ele))]))
icgp.ranges <- range(icgp)
icgp <- cut(icgp, 10, labels=colorRampPalette(c("green", "black", "red"))(10))
icgp.image <- as.raster(matrix(rev(levels(icgp)), ncol=1))
icgp <- as.character(icgp)

plotMotifStackWithRadialPhylog(phylog=phylog, pfms=sig, 
                               col.bg=methods, col.bg.alpha=.3, 
                               col.inner.label.circle=gpCol, 
                               inner.label.circle.width=0.02, 
                               labels.leaves=leaveNames, cleaves=.1, 
                               circle=.95, circle.motif=1.8, clabel.leaves=.8,
                               motifScale="logarithmic", angle=358, 
                               plotIndex=TRUE, IndexCex=.6, groupDistance=gpDis)

legend(-2, 2.2, legend=c("Mm/Uniprobe/PBM/Seed-And-Wobble", 
                         "Mm/CIS-BP/PBM/DREAM5", "Mm/CIS-BP/SELEX"), 
       fill=highlightCol(colorSet[c("Mm", "MmDREAM", "Ms")], .3), 
       border="white", lty=NULL, bty = "n")
```

```{r MmPBM_SELEX.ALL, fig.width=12, fig.height=12, dpi=72, fig.cap=fn("Alignment of mouse motifs from different methods", "E.b"), echo=FALSE}
motifCircos(phylog=phylog, pfms=motifs, col.tree.bg=methods, 
            col.tree.bg.alpha=.3, r.rings=c(.1, .1), 
            col.rings=list(icgp, rep("white", length(icgp))), 
            col.inner.label.circle=gpCol, inner.label.circle.width=0.02, 
            labels.leaves=leaveNames, cleaves=.1, r.tree=1, r.leaves=.5, 
            clabel.leaves=.9, motifScale="logarithmic", angle=358, 
            plotIndex=TRUE, IndexCex=.6, groupDistance=gpDis)

legend(-2.4, 2.4, legend=c("Mm/Uniprobe/PBM/Seed-And-Wobble", 
                           "Mm/CIS-BP/PBM/DREAM5", "Mm/CIS-BP/SELEX"), 
       fill=highlightCol(colorSet[c("Mm", "MmDREAM", "Ms")], .3), 
       border="white", lty=NULL, bty = "n")

text(2.05, 2.3, labels="Information Content")
rasterImage(icgp.image, 2, 1.8, 2.04, 2.2, interpolate = FALSE)
text(c(2.05, 2.05), c(1.8, 2.2), labels=format(icgp.ranges, digits=3), adj=0)

detach(MmPBM_SELEX)
```

# Supplementary Figure 4F. Align motifs from mouse generated by different platforms (SELEX, PBM-SW and PBM-dream5) using MotIV
```{r MmPBM_SELE.MotIV, fig.width=12, fig.height=12, dpi=72, fig.cap=fn("Alignment of mouse data generated by different methods using MotIV", "F.a"), echo=FALSE}
gpDis <- 0.005 
MmPBM_SELEX.MotIV <- getMotIVOut("pcmsDatasetMmPBM_SELEX", 
                                 cc="ALLR", align="SWU",
                                 groupDistance=gpDis)
attach(MmPBM_SELEX.MotIV)

##set methods color
methods <- factor(grepl("_SELEX", leaves) + grepl("_M\\d{4}", leaves))
levels(methods) <- c("Mm", "MmDREAM", "Ms")
##c("PBM/Uniprobe/Seed-And-Wobble", "PBM/CIS-BP/DREAM5", "SELEX/CIS-BP")
levels(methods) <- colorSet[levels(methods)]
methods <- as.character(methods)

leaveNames <- gsub("_(SELEX|PBM)", "", gsub("_(Jolma|Berger)", "", leaves))

## calculate average of top 8 position information content for each motif
icgp <- sapply(sapply(unaligned.pfms, getIC), function(.ele) 
  mean(sort(.ele, decreasing=TRUE)[1:min(8, length(.ele))]))
icgp.ranges <- range(icgp)
icgp <- cut(icgp, 10, labels=colorRampPalette(c("green", "black", "red"))(10))
icgp.image <- as.raster(matrix(rev(levels(icgp)), ncol=1))
icgp <- as.character(icgp)

plotMotifStackWithRadialPhylog(phylog=phylog, pfms=sig, 
                               col.bg=methods, col.bg.alpha=.3, 
                               col.inner.label.circle=gpCol, 
                               inner.label.circle.width=0.02, 
                               labels.leaves=leaveNames, cleaves=.1, 
                               circle=.95, circle.motif=1.8, clabel.leaves=.8,
                               motifScale="logarithmic", angle=358, 
                               plotIndex=TRUE, IndexCex=.6, groupDistance=gpDis)

legend(-2, 2.2, legend=c("Mm/Uniprobe/PBM/Seed-And-Wobble", 
                         "Mm/CIS-BP/PBM/DREAM5", "Mm/CIS-BP/SELEX"), 
       fill=highlightCol(colorSet[c("Mm", "MmDREAM", "Ms")], .3), 
       border="white", lty=NULL, bty = "n")

```


```{r MmPBM_SELEX.MotIV.ALL, fig.width=12, fig.height=12, dpi=72, fig.cap=fn("Alignment of mouse motifs generated by different methods using MotIV", "F.b"), echo=FALSE}
motifCircos(phylog=phylog, pfms=motifs, col.tree.bg=methods, 
            col.tree.bg.alpha=.3, r.rings=c(.1, .1), 
            col.rings=list(icgp, rep("white", length(icgp))), 
            col.inner.label.circle=gpCol, inner.label.circle.width=0.02, 
            labels.leaves=leaveNames, cleaves=.1, r.tree=1, r.leaves=.5, 
            clabel.leaves=.9, motifScale="logarithmic", angle=358, 
            plotIndex=TRUE, IndexCex=.6, groupDistance=gpDis)

legend(-2.4, 2.4, legend=c("Mm/Uniprobe/PBM/Seed-And-Wobble", 
                           "Mm/CIS-BP/PBM/DREAM5", "Mm/CIS-BP/SELEX"), 
       fill=highlightCol(colorSet[c("Mm", "MmDREAM", "Ms")], .3), 
       border="white", lty=NULL, bty = "n")

text(2.05, 2.3, labels="Information Content")
rasterImage(icgp.image, 2, 1.8, 2.04, 2.2, interpolate = FALSE)
text(c(2.05, 2.05), c(1.8, 2.2), labels=format(icgp.ranges, digits=3), adj=0)

detach(MmPBM_SELEX.MotIV)
```

# Supplementary Figure 4G. Align motifs of engrailed orthologs using MatAlign
```{r EN, fig.width=8, fig.height=4, dpi=72, fig.cap=fn("Alignment of engrailed orthologs using MatAlign", "G"), echo=FALSE}
gpDis <- 2
EN <- getMatAlignOut("pcmsDatasetEn", groupDistance=gpDis)
attach(EN)

##set methods color
species <- gsub("^(Dm|Hs|Mm).*?$", "\\1", leaves)
species[grepl("_M\\d+", leaves) & species=="Mm"] <- "MmDREAM"
species <- colorSet[species] ##c("Dm/FFS/B1H", "Hs/CIS-BP/SELEX", "Mm/Uniprobe/PBM/\nSeed-And-Wobble")

leaveNames <- gsub("^(Dm|Mm|Hs)_", "", leaves)

motifPiles(phylog=phylog, motifs, r.tree=.5, sig, 
           col.tree=species, col.leaves=species, 
           col.pfms2=gpCol, col.pfms2.width=.01, 
           labels.leaves=leaveNames, plotIndex=TRUE, 
           IndexCex=2, groupDistance=gpDis, clabel.leaves=2)
legend("topleft", legend=c("Dm/FFS/B1H", "Hs/CIS-BP/SELEX", 
                           "Mm/Uniprobe/PBM/\nSeed-And-Wobble", 
                           "Mm/CIS-BP/PBM/Dream5"), 
       fill=colorSet[c("Dm", "Hs", "Mm", "MmDREAM")], 
       border="white", lty=NULL, bty = "n", cex=.5)

detach(EN)
```


# Supplementary Figure 4H. Align motifs of engrailed orthologs using MotIV 
```{r EN.MotIV, fig.width=8, fig.height=4, dpi=72, fig.cap=fn("Alignment of engrailed orthologs using MotIV", "H"), echo=FALSE}
gpDis <- 0.005 
EN.MotIV <- getMotIVOut("pcmsDatasetEn", 
                                    cc="ALLR", align="SWU", 
                                    groupDistance=gpDis)
attach(EN.MotIV)

##set methods color
species <- gsub("^(Dm|Hs|Mm).*?$", "\\1", leaves)
species[grepl("_M\\d+", leaves) & species=="Mm"] <- "MmDREAM"
##c("Dm/FFS/B1H", "Hs/CIS-BP/SELEX", "Mm/Uniprobe/PBM/\nSeed-And-Wobble")
species <- colorSet[species]

leaveNames <- gsub("^(Dm|Mm|Hs)_", "", leaves)

motifPiles(phylog=phylog, motifs, r.tree=.5, sig, 
           col.tree=species, col.leaves=species, 
           col.pfms2=gpCol, col.pfms2.width=.01, 
           labels.leaves=leaveNames, plotIndex=TRUE, 
           IndexCex=2, groupDistance=gpDis, clabel.leaves=2)
legend(0.1, 0.65, legend=c("Dm/FFS/B1H", "Hs/CIS-BP/SELEX", 
                           "Mm/Uniprobe/PBM/\nSeed-And-Wobble", 
                           "Mm/CIS-BP/PBM/Dream5"), 
       fill=colorSet[c("Dm", "Hs", "Mm", "MmDREAM")], 
       border="white", lty=NULL, bty = "n", cex=.5)

detach(EN.MotIV)
```

# Supplementary Figure 4I. Align motifs from fly, mouse and human using MatAlign
```{r FlyMousePBMHuman, fig.width=12, fig.height=12, dpi=72, fig.cap=fn("Alignment of motifs from fly, mouse and human using MatAlign", "I"), echo=FALSE}
gpDis <- 2.5
FlyMousePBMHuman <- getMatAlignOut("pcmsDatasetFlyMousePBMHuman",
                                   groupDistance=gpDis)
attach(FlyMousePBMHuman)

##set methods color
species <- factor(gsub("^(Dm|Mm|Hs).*$", "\\1", leaves))
##c("Dm/FFS/B1H", "Hs/CIS-BP/SELEX", "Mm/Uniprobe/PBM/Seed-And-Wobble")
levels(species) <- colorSet[c("Dm", "Hs", "Mm")] 
species <- as.character(species)

leaveNames <- gsub("^(Dm|Mm|Hs)_", "", leaves)

plotMotifStackWithRadialPhylog(phylog=phylog, pfms=sig, 
                               col.bg=species, col.bg.alpha=.3, 
                               col.inner.label.circle=gpCol, 
                               inner.label.circle.width=0.02, 
                               labels.leaves=leaveNames, 
                               cleaves=.1, circle=1.5, circle.motif=1.8, 
                               clabel.leaves=.5, motifScale="logarithmic", 
                               angle=358, plotIndex=TRUE, IndexCex=.6, 
                               groupDistance=gpDis)

legend(-2, 2.2, 
       legend=c("Dm/FFS/B1H", 
                "Mm/Uniprobe/PBM/Seed-And-Wobble", 
                "Hs/CIS-BP/SELEX"), 
       fill= highlightCol(colorSet[c("Dm", "Hs", "Mm")], alpha=.3), 
       border="white", lty=NULL, bty = "n")

detach(FlyMousePBMHuman)
```

# Supplementary Figure 4J. Align motifs from fly vs mouse and human using MotIV
```{r FlyMousePBMHuman.MotIV, fig.width=12, fig.height=12, dpi=72, fig.cap=fn("Alignment of motifs from fly, mouse and human using MotIV", "J"), echo=FALSE}
gpDis <- 0.005
FlyMousePBMHuman.MotIV <- getMotIVOut("pcmsDatasetFlyMousePBMHuman", 
                                      cc="ALLR", align="SWU", 
                                      groupDistance=gpDis)
attach(FlyMousePBMHuman.MotIV)

##set methods color
species <- factor(gsub("^(Dm|Mm|Hs).*$", "\\1", leaves))
##c("Dm/FFS/B1H", "Hs/CIS-BP/SELEX", "Mm/Uniprobe/PBM/Seed-And-Wobble")
levels(species) <- colorSet[c("Dm", "Hs", "Mm")] 
species <- as.character(species)

leaveNames <- gsub("^(Dm|Mm|Hs)_", "", leaves)

plotMotifStackWithRadialPhylog(phylog=phylog, pfms=sig, 
                               col.bg=species, col.bg.alpha=.3, 
                               col.inner.label.circle=gpCol, 
                               inner.label.circle.width=0.02, 
                               labels.leaves=leaveNames, 
                               cleaves=.1, circle=1.5, circle.motif=1.8, 
                               clabel.leaves=.5, motifScale="logarithmic", 
                               angle=358, plotIndex=TRUE, IndexCex=.6, 
                               groupDistance=gpDis)

legend(-2, 2.2, 
       legend=c("Dm/FFS/B1H", 
                "Mm/Uniprobe/PBM/Seed-And-Wobble", 
                "Hs/CIS-BP/SELEX"), 
       fill= highlightCol(colorSet[c("Dm", "Hs", "Mm")], alpha=.3), 
       border="white", lty=NULL, bty = "n")

detach(FlyMousePBMHuman.MotIV)
```

```{r}
sessionInfo()
```