---
title: "Supplementary Figure 6. Motif Clusters for HOX and Caudal Related TFs"
output: 
  pdf_document:
    fig_caption: yes
    toc: no
    includes:
      in_header: header.tex
  html_document:
    fig_caption: yes
    self_contained: no
    toc: no
---

```{r setup, include=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(progress=FALSE, verbose=FALSE, keep_md=FALSE)
opts_chunk$set(comment=NA, warning=FALSE, message=FALSE, fig.width=8, fig.pos="!htb")
```
```{r loadLib, message=FALSE, warning=FALSE, error=FALSE,echo=FALSE}
##load the library
library(motifStack)
```
```{r dataInput, echo=FALSE}
source("safeColor.R")
outpath <- "output"
matalign_path <- "./app/matalign-v4a"
neighbor_path <- "./app/neighbor.app/Contents/MacOS/neighbor"
MatAlign2tree_path <- "./MatAlign2tree.pl"
getMatAlignOut <- function(pcmpath, groupDistance=2.5, trim=0.2){
    pcms <- readPCM(pcmpath)
    pcms <- safeColor(pcms)
    pfms<-lapply(pcms,pcm2pfm)
    system(paste("perl", MatAlign2tree_path, "--in . --pcmpath", pcmpath, "--out", outpath,
             "--matalign", matalign_path, "--neighbor", neighbor_path, "--tree","UPGMA"))
    newickstrUPGMA <- readLines(con=file.path(outpath, "NJ.matalign.distMX.nwk"))
    phylog <- newick2phylog(newickstrUPGMA, FALSE)
    leaves <- names(phylog$leaves)
    motifs <- pfms[leaves]
    if(!is.na(groupDistance)){
        motifSig <- motifSignature(motifs, phylog, groupDistance=groupDistance,
                                   min.freq=1, trim=trim)
        sig <- signatures(motifSig)
        gpCol <- sigColor(motifSig)
    }else{
        motifSig <- NA
        sig <- NA
        gpCol <- NA
    }
    
    return(list(phylog=phylog, sig=sig, gpCol=gpCol, 
                motifs=DNAmotifAlignment(motifs), leaves=leaves, 
                unaligned.pfms=motifs))
}

pcmpath <- dir("pcmsUni", include.dirs=TRUE)[12]
matAlignOut <- lapply(file.path("pcmsUni", pcmpath), 
                      getMatAlignOut, groupDistance=4)
colorSet <- c("Dm"="#009E73", 
              "Mm"="#D55E00", "Ms"="#E69F00", 
              "Hs"="#CC79A7")

```

```{r expand, echo=FALSE}
## kexpand is function to extend knitr.
kexpand<-function(.ele, id, cap, figheight, des){
    text <- paste("{{des}}\n\n```{r {{cap}},fig.cap='{{cap}}',fig.height={{figheight}},echo=",
                  ifelse(id==1,"TRUE","FALSE"),
                  "}\n\n
leaveNames <- gsub('^(Dm|Mm|Ms|Hs)_', '', .ele$leaves)\n\n
species <- colorSet[gsub(\"^(Dm|Mm|Ms|Hs).*$\", \"\\\\1\", .ele$leaves)]\n\n
motifPiles(phylog=.ele$phylog, .ele$motifs, 
           .ele$sig, 
           col.tree=species, col.leaves=species, 
           col.pfms2=.ele$gpCol, 
           col.pfms2.width=.01, labels.leaves=leaveNames, 
           plotIndex=c(FALSE,TRUE), IndexCex=1.5, 
           groupDistance=4, clabel.leaves=3,
           groupDistanceLineCol='#D55E00')\n\n
```\n\n", sep="")
    cat(knit(text=knit_expand(text=text)))
}
```

```{r plots, results='asis', echo=FALSE}
sta <- mapply(function(.ele, name, ID, des){
    TF.name <- gsub("Platform", "", gsub("Cluster", "", name))
    #kexpand(.ele, id=ID, cap=paste0("Supplementary Figure 6.", TF.name),
    #  figheight=ceiling(.25*length(.ele$leaves)), des)
    kexpand(.ele, id="", cap="__Supplemental Figure 6__. Motifs of HOX and caudal related TFs are shown in purple (human), brown and orange (mouse) and green (fly). These TFs cluster into two different motif groups, one with fly motifs and the other without fly motifs",
        figheight=ceiling(.25*length(.ele$leaves)), "")    
}, matAlignOut, pcmpath, 1:length(pcmpath), "")

```